---
title: "Project: Simulation and R"
author: "A.V Hurtado Vanessa and Narain Ritish"
date: "28/03/2023"
output:
  html_document:
      toc: TRUE
  pdf_document: 
      
      toc: TRUE
---

```{r setup, include=FALSE}
library(FactoMineR)
library(utils)
library(stats)
library(ggiraphExtra)
library(measurements)
library(dplyr)
library(corrplot)
library(kableExtra)
library(dplyr)
library(factoextra)
library(car)
library(ellipse)
#library(devtools)
library(webshot)
library(tidyverse)
library(SimDesign)
library(ggplot2)
library(ISwR)
library(MASS)
library(Sleuth3)
library(tidyr)
library(leaps)
library(GGally)
library(lmtest)
library(car)
library(scales)
library(rgl)
library(plot3D)
library(gganimate)
library(gifski)
library(av)
library(base)
set.seed(123)
```


# The simulation problem 

La tortue est posée au point $(0,0)$.

 * A l'étape 1, la tortue se déplace de $+u$ avec $u$ tiré au hasard uniformément parmi $(0,1),(1,0),(0,-1),(-1,0)$

 * A l'étape elle se redéplace par le même procédé aléatoire.

 * On répète ce déplacement $n$ fois.

 * On note $N_{n}$ le nombre de fois où la tortue revient à un point déjà visité dans le passé.

 * Pour $n=100$, puis 1000 puis 10000 , utilisez la méthode de monté Carlo pour avoir une idée graphique de la loi de $N_{n}$.


$\textbf{Position and movement}$

$N_n$ is the number of times the turtle revisits a point it has already visited in the first $n$ steps of its random walk.
The turtle starts at a point $(x_0, y_0)$ on a two-dimensional grid and it moves to a neighbourhood point chosen uniformly at random from the set $\{(x_0+1, y0), (x_0-1, y0), (x_0, y_0+1), (x_0, y_0-1)\}$ at each step. Let $(x_i, y_i)$ be the position of the turtle after i steps. Then, Nn is defined as:

$$N_n = \#\{i, j | 0 \leq i < j \leq {n-1} \& (x_i, y_i) = (x_j, y_j)\}$$

Therefore, $N_n$ is a random variable that depends on the random walk of the turtle and its distribution gives us information about the probability of the turtle revisiting points in the first $n$ steps of its walk. 

We will present the random walk algorithm in 1 dimension. Here the turtle is at position $0$. It can move to ${-1,1}$ after 1 step and move to either $-2$ or $0$ if the first step was $-1$.


## 1-D random walk algorithm {.tabset}

\textbf{1-D random walk algorithm 1}

```{r,echo=TRUE}
no_steps <- 1000
number_Walks <- 300

# We create a matrix to store the position of each step
positions <- matrix(0, ncol = no_steps + 1, nrow = number_Walks)

# We create a loop to calculate the position of each step

for (r in 1:number_Walks)
{
u <- 0 # The initial position at u= 0

for (i in 1:no_steps) 
{
step <- runif(1, -1, 1) # Generate a uniform random number between -1 and 1
u <- u + step
positions[r, i + 1] <- u # The new position of 1 random walk
}
}
```

\textbf{1-D random walk algorithm 2}

```{r, echo=TRUE}
ran_walk_1D2 <-function(number_Walks,no_steps)
{
# We create a matrix of uniform generated variables U([-1,1])
mat_grid <- matrix(runif(number_Walks * no_steps, -1, 1), ncol = no_steps)

# We calculate the position of each steps here
positions2 <- apply(cbind(rep(0, number_Walks), mat_grid), MARGIN = 2, FUN = cumsum)

return(positions2)
}
```

```{r, echo=TRUE}
simul1d_1 <- ran_walk_1D2(number_Walks = 1, no_steps = 100)
simul1d_100 <- ran_walk_1D2(number_Walks = 100, no_steps = 100)
simul1d_1000 <- ran_walk_1D2(number_Walks = 100, no_steps = 1000)
simul1d_10000 <- ran_walk_1D2(number_Walks = 100, no_steps = 10000)
```


### Plot of the 1-D random walk 

```{r,echo=TRUE}
print("A random walk with 100 simulations")
plot(simul1d_1, type = "l", xlab = "Steps", ylab = "Position", main = "Random Walk Path", col = "dark green", fg = "dark green")
```

### Multiple 1-D random walk 

```{r, echo=TRUE}
matplot(simul1d_100, type = "l", col = 1:number_Walks, xlab = "Number of steps", ylab = "Position")
matplot(simul1d_1000, type = "l", col = 1:number_Walks, xlab = "Number of steps", ylab = "Position")
matplot(simul1d_10000, type = "l", col = 1:number_Walks, xlab = "Number of steps", ylab = "Position")
```

### Probability distribution generated from multiple random walks

```{r,echo=TRUE}
hist_100 <- hist(simul1d_100, breaks = 80, freq = FALSE,col = "dark green", fg = "dark green",
                  xlab = "Number of steps", ylab = "Length of random walk",
                   main = "Distribution of the random variables generated")

hist_1000 <- hist(simul1d_1000, breaks = 80, freq = FALSE,col = "dark green", fg = "dark green",
                  xlab = "Number of steps", ylab = "Length of random walk",
                   main = "Distribution of the random variables generated")

hist_10000 <- hist(simul1d_10000, breaks = 80, freq = FALSE,col = "dark green", fg = "dark green",
                   xlab = "Number of steps", ylab = "Length of random walk",
                   main = "Distribution of the random variables generated")
```


```{r, echo=TRUE}
cdf_100 <- plot(ecdf(ran_walk_1D2(100,100)), main = "Empirical cumulative distribution with 100 steps ")
cdf_1000 <- plot(ecdf(ran_walk_1D2(100,1000)), main = "Empirical cumulative distribution with 1000 steps")
cdf_10000 <- plot(ecdf(ran_walk_1D2(100,10000)), main = "Empirical cumulative distribution 10000 steps")
```



From the graphical results show the histogram and the cumulative density function, we can see that the random walk converges closer to the normal distribution with a higher number of simulations $M$. Now, we will use monte carlo to have an idea of the distribution of the positions revisited from the turtle. We have considered the chi-squared distribution and poisson distribution for comparison of the distribution that will be generated. 



## 2-D random walk algorithm {.tabset}


\textbf{1-D random walk algorithm 1: Random walk only}
```{r}
Random_Walk_2D <- function(n_sim) 
  
{
pos_desp <- c(-1, 0, 1) # possible displacement values

n_dup <-numeric(n_sim) # initialize a vector to store the number of duplicated positions in each simulation

for (i in 1:n_sim) 
  {
 Alea <- sample(pos_desp, size = 2*n_sim, replace = TRUE) # generate a random sample of displacement values with replacement
 desp <- matrix(Alea, ncol = 2)                           # create a 2D matrix for the displacements with two columns, one for each                                                             dimension
 pos <- apply(X = desp, MARGIN = 2, FUN = cumsum)         # apply the cumsum function to the columns to get step N+1
  
return(pos)

}
}
```

\textbf{1-D random walk algorithm 2: Estimating the distribution of duplicate positions}
```{r}
set.seed(1234)
disp_2D_mc <- function(n_sim) 
  
{
  
pos_desp <- c(-1, 0, 1) # possible displacement values

n_dup <-numeric(n_sim) # initialize a vector to store the number of duplicated positions in each simulation

for (i in 1:n_sim) 
  {
 Alea <- sample(pos_desp, size = 2*n_sim, replace = TRUE) # generate a random sample of displacement values with replacement
 desp <- matrix(Alea, ncol = 2)                           # creates a 2D matrix for the displacements with two columns, one for each                                                             dimension
 pos <- apply(X = desp, MARGIN = 2, FUN = cumsum)         # apply the cumsum function to the columns to get step N+1
  
 n_dup[i] <- sum(duplicated(pos))                         # Counts the sum of duplicated positions
}
prob_dup <- sum(n_dup>0)/n_sim                            # calculate the probability of duplicated positions (N_n/n)

Num_dup <-  max(n_dup)
hist_dup <- hist(n_dup, main = paste("Histogram of duplicated points (Simulations =", n_sim, ")"), 
                                  xlab = "Number of duplicated points", ylab = "Frequency", breaks =n_sim/10,xlim = c(0,Num_dup),col = "dark green", fg = "dark green", )

E_X <- mean(n_dup)
x <- seq(0, max(n_dup))
Poiss <- dpois(x, E_X) #Density of poisson distribution
Chi2 <- dchisq(x, E_X) #Density of chi-square distribution

lines(x,Poiss*sum(n_dup), col = "red") 
lines(x,Chi2*sum(n_dup), col = "blue")
legend("bottomleft", legend = c("Simulated distribution", "Poisson distribution", "Chi Square distribution"), 
       lty = c(1, 1), col = c("black", "red", "blue"), cex = 0.7)

return(n_dup)
print("Histogram of the Simulated duplicated positions\n", hist_dup,
      "\n Number of duplicated positions of the turtle\n", Num_dup,
      ) 
      
}

```






### Probability distribution generated from 100 Simulations having duplicated positions
```{r, echo=TRUE}
#simul2d_100 <-disp_2D_mc(100)
simul2d_100.cdf <-plot(ecdf(disp_2D_mc(100)), main = "Empirical cumulative distribution 100 steps",col = "dark green", fg = "dark green")
```

### Probability distribution generated from 1000 Simulations having duplicated positions
```{r, echo=TRUE}
#simul2d_1000 <-disp_2D_mc(1000)
simul2d_1000.cdf <-plot(ecdf(disp_2D_mc(1000)), main = "Empirical cumulative distribution 1000 steps",col = "dark green", fg = "dark green")
```


### Probability distribution generated from 10000 Simulations having duplicated positions
```{r, echo=TRUE}
#simul2d_10000 <-disp_2D_mc(10000)
simul2d_10000.cdf <-plot(ecdf(disp_2D_mc(10000)), main = "Empirical cumulative distribution 10000 steps" )
```

We have successfully simulated the random walk in 2 dimension. From the same analysis in the 1 dimension results, we can see that the distribution of the positions duplicated from the Monte-Carlo simulation converges this time to a poisson distribution with the parameter $\lambda$ for $P_{o}(\lambda)$ as $n -> +\inf$ with the mean of the simulation. The result makes sense with the poisson distribution property as it represents occurrences for non regular events. 


# Random walk animations {.tabset}

This section is an additional piece of the code that animates the 2-D random walk with the function created \textbf{Random_Walk_2D}

##  100 Simulations
```{r, echo=TRUE, warning=FALSE, message=FALSE}
Random_2D_100 <- Random_Walk_2D(100)

# Plot of the path
plot(Random_2D_100, type = "l", col = "blue", xlab = "X", ylab = "Y", main = "Random Walk Path")


set.seed(123)

df1 <- data.frame(x = Random_2D_100[,1], y = Random_2D_100[,2])


p1 <- ggplot(df1, aes(x, y)) +
  geom_line(color = "blue") +
  xlab("X") +
  ylab("Y") +
  ggtitle("Random Walk Path")

animation1 <- p1 + transition_reveal(seq_along(df1$x)) +
  ease_aes('linear') +
  labs(title = "Step: {frame}")

# Plot + Animation
animate(animation1)


```

## Plot of the duplicated random walk

```{r echo=TRUE}
set.seed(123)
Random_2D_10000 <- Random_Walk_2D(10000)

plot(Random_2D_10000, type = "l", col ="green", xlab = "X Range", ylab = "Y Range", main = "Random Walk Path")

```
